---
title: '00'
output: html_document
---


```{r  include = FALSE}
# I added the cache = TRUE to try to speed up but not sure if this is right
## catch years as numeric or character (returned from R calls) and do not insert commas, e.g. "1991".
knitr::knit_hooks$set(inline = function(x) {
   ifelse(!(x %in% c(1891:2032)), prettyNum(x, big.mark=","),x)
 # ifelse(!(x %in% c(1891:2032)  | length(grep(paste(1891:2032,collapse = "|"),x)) == 0 ), prettyNum(x, big.mark=","),x)
})
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::knit_hooks$set(plot = function(x,options) {
      base = knitr::opts_knit$get('base.url')
      if (is.null(base)) base = ''
      alt = ifelse (is.null(options$alt),"",options$alt)
      cap = ifelse (is.null(options$caption),"",options$caption)
      if (alt != ""){
        sprintf('![%s](%s%s "%s")', cap, base, x, alt)
      } else {
        sprintf('![%s](%s%s)', cap, base, x)  
        }
  })
load("00opts.Rdata")
# changing what I used in the draft function
spp = 'sablefish'
Spp = 'Sablefish'



if(file.exists("00mod.Rdata")){
  load("00mod.Rdata") ## model
  load("benchmod.rdata") ## o19
} else {
  ## r4ss failed with the changes in May 7 caused by SPR. There's something wrong with exec summary
  ## in a time crunch so reverting r4ss
  # remotes::install_github('r4ss/r4ss@b0974690e1367a8b92cfaf7d9968cc9a697806ba') 
  ## extract base case into doc and save as model
  # model <- read_model(mod_loc = "C:/Users/mkapur/Dropbox/UW/assessments/sab-2021/council_runs/torc-lq-base-bnds-hess/", save_loc  = getwd())
  # save(model, file = here('00mod.Rdata'))
  
#   print("Model output not being read. Please run the read_model() function.")
}
# data_fig_loc = "//nwcfile/FRAM/Assessments/CurrentAssessments/DataModerate_2021/copper_rockfish/data/biology/plots"
```

```{r, echo = FALSE}
require(dplyr)
options(scipen = 999)

if(basename(here()) == 'sab-2021'){
  refpts <- read.csv(here('doc', 'tables','e_referencepoints_es.csv'))
  ssbes <- read.csv(here('doc',  'tables','b_ssb_es.csv'))
} else{
  refpts <- read.csv(here( 'tables','e_referencepoints_es.csv'))
  ssbes <- read.csv(here(  'tables','b_ssb_es.csv'))
}


derived_quants <- model$derived_quants %>% mutate(lci = Value - 1.96*StdDev,
                                                  uci = Value + 1.96*StdDev ) %>%
  mutate(Value = ifelse(Value > 100, round(Value), round(Value,3)),
         lci = ifelse(lci > 100, round(lci), round(lci,3)),
         uci = ifelse(uci > 100, round(uci), round(uci,3)))


cpueyearrange<- model$cpue %>% select(Fleet_name, Yr) %>%
  group_by(Fleet_name) %>%
  summarise(miny = min(Yr), maxy = max(Yr)) %>%
  mutate(yrrange = paste(miny,maxy, sep = "-")) %>%
  select(Fleet_name, yrrange) %>%
  tidyr::pivot_wider(., names_from = Fleet_name, values_from = yrrange)
```


\newcommand{\textrefpt}{Unfished spawning biomass was estimated to be `r round(derived_quants[derived_quants$Label == "SSB_unfished", "Value"])` mt (`r noquote(refpts$Interval[1])` $\sim$95\% interval). The abundance of sablefish was estimated to have declined to near the target during the period 1980-2000. The estimate of the target spawning biomass was `r derived_quants[derived_quants$Label == "SSB_Btgt", "Value"]` (`r derived_quants[derived_quants$Label == "SSB_Btgt", "lci"]`-`r derived_quants[derived_quants$Label == "SSB_Btgt", "uci"]`, $\sim$95\%  interval). The stock was estimated to be  `r ifelse(derived_quants[derived_quants$Label == paste0("SSB_", model$endyr+1), "Value"]<derived_quants[derived_quants$Label == "Dead_Catch_Btgt", "Value"],"below","above")` the target stock size in the beginning of `r model$endyr+1` at `r derived_quants[derived_quants$Label == paste0("SSB_", model$endyr+1), "Value"]` mt (`r derived_quants[derived_quants$Label == paste0("SSB_", model$endyr+1), "lci"]`-`r derived_quants[derived_quants$Label == paste0("SSB_", model$endyr+1), "uci"]`, $\sim$95\% interval). The stock was estimated to be `r ifelse(derived_quants[derived_quants$Label == "Bratio_2020", "Value"]<derived_quants[derived_quants$Label == "B_MSY/SSB_unfished", "Value"],'below','above')` the depletion level that would lead to maximum yield (0.4) (Figures \ref{fig:yield1} and \ref{fig:SPR4_phase}). The estimate of the stock's current `r model$endyr+1` level of depletion was `r derived_quants[derived_quants$Label == "Bratio_2021", "Value"]`.}

\newcommand{\textexplstatus}{ Although the estimated productivity and absolute scale of the stock are poorly informed by the available data and are, therefore, sensitive to changes in model structure and treatment of data, all sensitivity or alternate models evaluated showed a declining trend in biomass since the 1970s followed by a recent increase in biomass (Figures \ref{fig:SPR3} and \ref{fig:ts_summaryF}). The spawner potential ratio ($SPR$) relative to the fishing mortality target or overfishing level ($SPR_{45\%}$) that stabilizes the stock at the target (reported as $(1-SPR)/[1-SPR_{45\%}]$), was greater than 1 (thus exceeding the target rate) during nearly half of the years from 1976 through 2000, has been below the target since, and was between 0.62 and 0.76 from 2015-2019, descending to 0.40 in 2020.
'Relative 1-SPR' in Table \ref{tab:exploitES} refers to $(1-SPR)/[1-SPR_{45\%}]$ ; where 1 is the target exploitation rate, and values over 1 indicate overexploitation relative to this proxy. While highly uncertain, the
absolute equilibrium yield at the estimated fishing mortality that leads to the maximum sustainable yield ($F_{MSY}$) is `r round(derived_quants[derived_quants$Label == "Dead_Catch_MSY", "Value"])` mt (`r round(derived_quants[derived_quants$Label == "Dead_Catch_MSY", "lci"])`-`r round(derived_quants[derived_quants$Label == "Dead_Catch_MSY", "uci"])`, $\sim$95\% interval), while the proxy SPR rate of 0.45 leads to a proxy MSY of 8,350 mt (3,924 - 12,777, ~95\% interval).}

\newcommand{\textsciuncertainty}{The time series of multiplicative buffer fractions that are a function of $P^*$ and the time series of $\sigma$s provide the multipliers on the overfishing limit; these values are all less than 1 for category 1 stocks.  $\sigma$ for sablefish is the time-varying category 1 value, which starts at 0.5 in the year after the (update) assessment and increases throughout the projection period. The uncertainty around the OFL value for the first forecast year (2022) is `r round(sqrt(log((model$derived_quants[model$derived_quants$Label == 'OFLCatch_2022','StdDev']/model$derived_quants[model$derived_quants$Label == 'OFLCatch_2022','Value'])^2+1)),3)`; the uncertainty around spawning output in that same year is 0.298, both less than 0.5.
}


\newcommand{\textprojections}{Previous sablefish stock assessments have been designated as Category 1 stock assessments. Projections and decision tables are based on $P^*$=0.45, the adopted value for the most recent management cycle, and the values of $\sigma$ adopted by the Pacific Fisheries Management Council for stock projections. The time series of multiplicative buffer fractions that are a function of $P^*$ and the time series of $\sigma$s provide the multipliers on the overfishing limit; these values are all less than 1 for category 1 stocks. $\sigma$ for sablefish is the time-varying category 1 value, which starts at 0.5 in the year after the (update) assessment and increases throughout the projection period.  The uncertainty around the OFL value for the first forecast year (2022) is `r round(sqrt(log((model$derived_quants[model$derived_quants$Label == 'OFLCatch_2022','StdDev']/model$derived_quants[model$derived_quants$Label == 'OFLCatch_2022','Value'])^2+1)),3)`; the uncertainty around spawning output in that same year is 0.298, both less than 0.5. The multipliers are combined with OFLs to calculate the ABC values. The Council sets ACL values which cannot exceed (with limited exceptions) the ABCs as modified by the 40-10 rule. The total catches in `r model$endyr+1` and `r model$endyr+2` were set at the Pacific Fisheries Management Council Groundfish Management Team requested values, below the Pacific Fisheries Management Council annual catch limits for sablefish. The average ratio between GMT-specified 2021-2022 catches were used to distribute catches among the fisheries for forecasted years.

Projections are provided through `r model$nforecastyears+model$endyr` (Table \ref{tab:projectionES}). Current medium-term projections from the base model under the Pacific Fisheries Management Council 40-10 harvest control rule estimate that the stock will remain above the target stock size of 40\% of the estimated unfished spawning biomass during the projection period.  Forecasts from the `r o19$endyr+1` benchmark assessment projected the spawning biomass to increase by `r paste0(round((o19$timeseries$SpawnBio[o19$timeseries$Yr == model$endyr+1]-o19$timeseries$SpawnBio[o19$timeseries$Yr == model$endyr-3])/o19$timeseries$SpawnBio[o19$timeseries$Yr == model$endyr-3]*100))`\% from `r model$endyr-3` to `r model$endyr+1` given specified harvests, whereas the current assessment estimated the increase at `r paste0(round((model$timeseries$SpawnBio[model$timeseries$Yr == model$endyr+1]-model$timeseries$SpawnBio[model$timeseries$Yr == model$endyr-3])/model$timeseries$SpawnBio[model$timeseries$Yr == model$endyr-3]*100))`\%. The estimate of unexploited spawning biomass (in the year of each assessment) is `r paste0(round((model$derived_quants[grep('SSB_Virgin',model$derived_quants$Label),'Value']-o19$derived_quants[grep('SSB_Virgin',o19$derived_quants$Label),'Value'])/model$derived_quants[grep('SSB_Virgin',model$derived_quants$Label),'Value']*100))`\% higher than that estimated in `r o19$endyr+1` and 19\% lower than the 2011 estimate. Relative unfished biomass in `r model$endyr+1` was estimated at `r round(model$derived_quants[grep('Bratio_2021',model$derived_quants$Label),'Value'],2)`, while the `r o19$endyr+1` benchmark assessment forecasted it to be  `r round(o19$derived_quants[grep('Bratio_2021',o19$derived_quants$Label),'Value'],2)`.}

\newcommand{\changesmade}{During the exploration of recent data for this update assessment, modelers identified increased discarding in the trawl fleet, for which the discard ratio nearly quadrupled between 2018 and 2019 (Figure \ref{fig:discard_fit}). In the first iteration of this update model, retention curve parameters were fixed, as discard length compositions were not included due to conflicts between the age and length data found in the 2019 benchmark assessment. Absent the data or structural flexibility to account for increased discarding, a model that conformed to the Terms of Reference (TOR) for an update assessment was unable to satisfactorily fit to the age composition data from the trawl fleets (Figure \ref{fig:torafleet3agecomps}) nor the WCGBT survey length composition data (Figure \ref{fig:torafleet8lencomps}), and greatly overestimated the 2019 index (Figure \ref{fig:torasurv}). Because the TOR model estimates retention for both fisheries in a single timeblock from 2011 onward, the discard data forced the model to generate many small fish, thus overestimating the model-expected index of abundance, the frequency of young and/or small individuals, and distorting the recruitment pattern (Figure \ref{fig:torapanel}).

We rectify the lack of fit to the data found in the TOR model by re-introducing the discard length compositions and time-blocking the retention curve to include a new block for the final two years of the model period (2019-2020; the benchmark model's terminal period for retention selectivity ran from 2011-2017). This adjustment resolved the aforementioned model fit issues (Figures \ref{fig:torcfleet8lencomps}-\ref{fig:basepanel}), and is herein presented as the "base model".}


\newcommand{\changesmadeshort}{During the exploration of recent data for this update assessment, modelers identified increased discarding in the trawl fleet, for which the discard ratio nearly quadrupled between 2018 and 2019 (Figure \ref{fig:discard_fit}). Absent the data or structural flexibility to account for increased discarding, a model that conformed to the Terms of Reference (TOR) for an update assessment was unable to satisfactorily fit to the age composition data from the trawl fleets (Figure \ref{fig:torafleet3agecomps}) nor the WCGBT survey length composition data (Figure \ref{fig:torafleet8lencomps}), and greatly overestimated the 2019 index (Figure \ref{fig:torasurv}). 
We rectify the lack of fit to the data found in the TOR model by re-introducing the discard length compositions and time-blocking the retention curve to include a new block for the final two years of the model period (2019-2020; the benchmark model's terminal period for retention selectivity ran from 2011-2017). This adjustment resolved the aforementioned model fit issues (Figures   \ref{fig:torcfleet8lencomps}-\ref{fig:basepanel}). The proposed base model presented here otherwise estimates parameters in the manner done in 2019, with the same priors.}

