## Base Model Results
## Convergence status

To test for convergence, 100 trials of the base model were ran using randomly generated alternative initial values for each estimated parameter. A value of 0.1 was used to define the uniform distribution that is transformed into cumulative normal space and subsequently used to calculate these initial values based on the parameter bounds. Thus, each trial perturbs the initial values used for minimization with the intention of causing the search to traverse a broader region of the likelihood surface (@MethotWetzel2013). The 100 iterations of the jitter test for the base model resulted in 18 model runs that failed to converge, 24 model runs that converged at or close to the total likelihood estimate value of the base case model run, and 58 model runs with total likelihood values higher than the base. This demonstrates that the jittered model was sensitive to the initial values of the parameters. The specification of both bounds and priors on individual parameters, together with penalties, weights on associated likelihoods, and high correlations among parameters can all affect jitter convergence. None of the trial runs were used to replace the base model.


```{r basecaseresults, echo = FALSE}
mf <- model$parameters[model$parameters$Label == "NatM_p_1_Fem_GP_1", ]
mm <- model$parameters[model$parameters$Label == "NatM_p_1_Mal_GP_1", ]
omf <- o19$parameters[o19$parameters$Label == "NatM_p_1_Fem_GP_1", ]
omm <- o19$parameters[o19$parameters$Label == "NatM_p_1_Mal_GP_1", ]

lmaxf <- model$parameters[model$parameters$Label == "L_at_Amax_Fem_GP_1", ]
lmaxm <- model$parameters[model$parameters$Label == "L_at_Amax_Mal_GP_1", ]
```

The biological parameters (growth and $M$) estimated using the base model and alternate models were reasonable. Growth parameters were consistent with those from previous `r spp` stock assessments and commensurate with the raw data (Table \ref{tab:basepars}). Female and male `r spp` showed similar rapid growth trajectories; with females growing to a slightly larger size at age 30 (`r round(lmaxf$Init,2)` cm) than males (`r round(lmaxm$Init,2)` cm) and showing a broader distribution of length at a given age (Figure \ref{fig:bio_53}). $M$ for females (`r signif(mf$Init,2)`) and males (0.60) were similar to values estimated in previous assessments (2011: 0.080 and 0.065 respectively; 2015: 0.076 and 0.062, 2019: `r signif(omf$Init,2)` and  `r signif(omm$Init,2)`, respectively; Figure \ref{fig:naturalmortalityprior}).

\begin{comment}
This assessment did not include time-varying growth. Differences were seen in the estimated weight-at-age compared to empirical weight at age collected by the WCGBT Survey (Figure \ref{fig:weightatageWCGBT}). These differences were more prominent in the most recent years, which might be a cohort effect. Future research could investigate methods for modeling time varying growth. \end{comment}

```{r echo = FALSE}
fixret <- model$parameters[model$parameters$Label == "Retain_L_infl_FIX(1)", ]
twlret <- model$parameters[model$parameters$Label == "Retain_L_width_TWL(3)", ]
q7 <- model$parameters[model$parameters$Label == "Q_extraSD_NWSLP(7)", ]
q6 <- model$parameters[model$parameters$Label == "Q_extraSD_AKSLP(6)", ]
q5 <- model$parameters[model$parameters$Label == "Q_extraSD_AKSHLF(5)", ]
c5 <- model$parameters[model$parameters$Label == "LnQ_base_AKSHLF(5)", ]
index5 <- model$cpue[model$cpue$Fleet_name == "AKSHLF", "Eff_Q"][1]
index6 <- model$cpue[model$cpue$Fleet_name == "AKSLP", "Eff_Q"][1]
index7 <- model$cpue[model$cpue$Fleet_name == "NWSLP", "Eff_Q"][1]
index8 <- model$cpue[model$cpue$Fleet_name == "NWCBO", "Eff_Q"][1]
```

Estimated selectivity curves for the trawl surveys varied, with the surveys that sample the continental slope sampling the broadest demographic of the `r spp` population and the Triennial Shelf Survey the most limited (Tables \ref{tab:basepars} and \ref{tab:selparest}; Figure \ref{fig:selectivityatageALL}). \begin{comment}The proportion of the spawning output that is unavailable to the surveys and fleets, which are all modeled using dome-shaped selectivity, has slightly decreased over time (Figure \ref{fig:unavailablessb}).\end{comment} The fixed gear fisheries showed males were less selected than females, and that individuals of approximately age 20 and older were much less available to the fishery on a relative basis (Figure \ref{fig:selectivityatageTWL}). The trawl fishery selected younger fish than the fixed gear fleet and showed little difference between males and females (Figure \ref{fig:selectivityatageTWL}). Retention schedules (Table \ref{tab:selparest}) showed rapidly increasing retention of age-1 fish for the fixed gear fishery but less than full retention of the largest individuals, likely due to some trip-limit based discarding or depredation of large fish during gear retrieval (Figure \ref{fig:selectivityatageFIX}). Full retention of the largest individuals was assumed since the beginning of the 2011 catch-shares program for the trawl fishery, with an increase in the minimum retention size for both sexes in the final two modeled years (Figure \ref{fig:twltimevaryslx}).

The base model fit the trend (decline, then stabilization, and increase) in the WCGBT Survey well (Figure \ref{fig:ind_NWFSC}), such that the added variance parameter was set to zero as was done in 2019. Fits to the NWFSC Slope Survey were generally flat (Figure \ref{fig:ind_NWFSCslope}), as might be expected for such short time-series. Fits to the AFSC Slope Survey suggest a decreasing trend during the late 1990s followed by and increase into the early 2000s (Figure \ref{fig:ind_AFSCslope}). Estimates of added variance were `r round(model$parameters[model$parameters$Label =="Q_extraSD_NWSLP(7)","Value"],2)` and `r round(model$parameters[model$parameters$Label =="Q_extraSD_AKSLP(6)","Value"],2)`, respectively (Table \ref{tab:selparest}). Given the time change in the estimate of $q$ for the Triennial Shelf Survey beginning in 1995, predicted survey values were also relatively flat over this period until the last two years of the survey (Figure \ref{fig:ind_AFSC}), although the estimated extra variance of `r round(q5$Value,2)` suggested a relatively poor fit to these data compared to other surveys. 

The fit to the sea-level index of recruitment was noisy, as expected, due to the relatively weak but persistent sea-level recruitment relationship, showing periods where the model was able to fit the data well, as well as periods with a lack of fit (Figure \ref{fig:ind_ENV}). The estimated added standard deviation was `r  round(as.numeric(model$index_variance_tuning_check$New_VarAdj[model$index_variance_tuning_check$Fleet == 4]),2)`, thus the sea-level index provided limited information regarding historical recruitment during model periods without other data.

The base model fit the length distributions from the WCGBT Survey well given that selectivity was modeled as age based, with residual patterns (Figures \ref{fig:torafleet8lencomps} and \ref{fig:lcomp_resid_8}) primarily generated through small mismatches in the model structure, likely due differences in growth, environmental conditions, or timing rather than misspecification of year-classes. The fits to the WCGBT Survey conditional-age-at-length distributions were good (Figures \ref{fig:acomp_resid_caal1}-\ref{fig:acomp_resid_caal4}). The slope survey fits to the marginal-age distributions also showed no glaring residual patterns in the age data (Figures \ref{fig:acomp_resid_1}- \ref{fig:acomp_resid_6}). The selection of younger `r spp` was evident for the Triennial Shelf Survey, with a larger residuals from 1995 forward (Figure \ref{fig:acomp_resid_5}).

Fits to the marginal age compositions for the fisheries were good (Figure \ref{fig:acaggregate}). All fisheries show relatively small residuals, with patterns of large cohorts moving through the population at some point (Figures \ref{fig:acomp_resid_1} and \ref{fig:acomp_resid_3}). Residual patterns might partially be the result of spatial differences in fishing, growth or movement. \begin{comment}As requested by the STAR panel, spatially explicit composition data north and south of $36^{\circ}$ N lat is provided in Appendix \ref{app:comp}.\end{comment}

The model was able to fit the mean body weights of the fishery discards and discard fractions well (Figures \ref{fig:bodywtfit} and \ref{fig:discard_fit}).

```{r, echo = FALSE}
bias <- model$breakpoints_for_bias_adjustment_ramp
biasmax <- bias$max_bias_adj
aa <- subset(model$derived_quants, Label == paste0("SSB_", model$endyr+1))
bb <- subset(model$derived_quants, Label == paste0("Bratio_", model$endyr+1))
```
Deviations about the estimated stock-recruitment function generally had high uncertainty prior to the mid-1970s, when the age-composition data first become informative about cohort strengths (Figure \ref{fig:recdevs_bias}). This stock assessment update was able to estimate cohort strengths further back in time due to the increased plus group, extended to 50 years (note that the data plus group is 70 years, whereas the modeled plus group is 50). The NWFSC and AFSC Slope Surveys, as well as the WCGBT Survey, all catch older fish that provided some information with respect to recruitment prior to the mid-1970s (the informative period for recruitment in past assessments). Including the sea level as a survey index of recruitment strength informs recruitment estimates in a limited fashion prior to the mid-1970s. The recruitment bias adjustment was set as recommended by @MethotTaylor2011.

`r Spp` recruitment was estimated to be highly variable with large amounts of uncertainty in individual recruitment events. Within this variability, there were sets of years with recruitment estimated consistently higher or lower than the long term mean (Figure \ref{fig:ts11}), with both the lowest and highest estimates occurring during the past 20 years.  Given a relatively high degree of recruitment variability, the estimated stock-recruitment function predicted a wide range of cohort sizes over the observed range of spawning biomass (Figure \ref{fig:SR_curve}).

Catches were input from the beginning of the time series (Table \ref{tab:allcatches}). The estimates of uncertainty around the point estimate of unfished biomass are large across the range of models explored within this assessment, suggesting that the unfished spawning biomass could range from just under 100,000 mt to over 200,000 mt. This uncertainty is largely due to the confounding of natural mortality, absolute stock size, and productivity. The point estimate of `r model$endyr` spawning biomass from the base model is `r round(aa$Value)` mt; however, the $\sim$ 95\% interval ranges broadly from `r round(aa$Value - 1.96 * aa$StdDev)` to `r round(aa$Value + 1.96 * aa$StdDev)` mt. The relative trend in spawning biomass is robust to uncertainty in the leading model parameters. The `r model$endyr+1` point estimate of spawning stock biomass is `r round(bb$Value,2) * 100`\% of the unfished state (approximate 95% interval: `r paste0(round(bb$Value - 1.96 * bb$StdDev,2)*100,"%")` to `r paste0(round(bb$Value +1.96 * bb$StdDev,2)*100,"%")`). Estimates indicate that the spawning biomass was near the target (Figure \ref{fig:ts9}). The estimated time-series of total, age-4+ (Figure \ref{fig:ts1_4}), and spawning biomass (Figure \ref{fig:ts7}) track one another closely (Table \ref{tab:timeseries}). 

\begin{comment}
Forecasts from the 2015 assessment update projected the spawning biomass to increase by 9.3% from 2015 to 2019 given specified harvests, whereas the current assessment estimated the increase at \Sexpr{pnt(abs(spbdecline), sigdiffs = 1)}%. Estimates of unexploited spawning biomass are \Sexpr{pnt(ssb0belowlastyear[1], 0)}% lower than that estimated in 2015 and \Sexpr{pnt(ssb0belowlastyear[2], 0)}% lower than the 2011 estimate. Percent of unfished biomass in 2019 was estimated at \Sexpr{pnt(depl, 0)}%, while the 2015 stock assessment forecasted it to be 38%.


## Data weighting{#dw}
```{r echo = FALSE, eval = FALSE}
fleetnames <- as.character(factor(unique(model$catch$Fleet_Name), 
  labels = c("Fixed-gear", "Pot", "Trawl")))[c(1, 3)]

# effn <- model$Dirichlet_Multinomial_pars
# effn$Type <- rep(c("length", "age"), each = max(model$fleet_ID) - 1)
# effn$Fleet <- c(
#   rownames(model$len_data_info)[model$len_data_info$CompError == 1],
#   rownames(model$age_data_info)[model$age_data_info$CompError == 1])
# effn$Fleet <- gsub("NWCBO", "WCGBT", effn$Fleet)
# effn$Fleet <- gsub("AKSHLF", "Triennial", effn$Fleet)
# colnames(effn) <- gsub("\\Theta", "$\\Theta$", colnames(effn))

varadj <- getvars(file.path(dir.basemodel, "control.ss_new"))
varadj <- varadj[varadj[, "fleet"] %in% c(1, 3), ]
discardmnad <- tapply(model$discard$Std_use, model$discard$Fleet, mean)
discardrmse <- tapply(model$discard$Dev, model$discard$Fleet,
  function(x) sqrt(sum(x^2) / length(x)))

meanbdwtmnad <- tapply(model$mnwgt$CV, model$mnwgt$Fleet, mean)
meanbdwtrmse <- tapply(model$mnwgt$Dev, model$mnwgt$Fleet,
  function(x) sqrt(sum(x^2) / length(x)))

fleetvar <- data.frame(
  "Fleet" = c("\\emph{Discard ratio}:", fleetnames,
              "\\emph{Mean body weight}:", fleetnames),
  "SD adj." = c(
    NA, varadj[varadj[, "type"] == 2, "oldvar"], 
    NA, varadj[varadj[, "type"] == 3, "oldvar"]),
  "Mean input SD (log) after adj." = c(NA, discardmnad, NA, meanbdwtmnad),
  "RMSE" = c(NA, discardrmse, NA, meanbdwtrmse))
colnames(fleetvar) <- c("Fleet", "SD adj.", "Mean SD after adj.", "RMSE")
```
\end{comment}

## Uncertainty and Sensitivity Analyses
Sensitivity analyses were performed to determine the sensitivity of the model results to a range of alternative assumptions. While the recent stock trend and estimates of stock depletion were similar among model sensitivities, a common theme is that the size of the unfished spawning biomass was highly uncertain. The available data for `r spp` were largely uninformative about the absolute stock size and productivity. This stock assessment model, given the data, was unable to discriminate between a larger, less productive stock and a smaller more productive stock, or many combinations in between. This could be due to the largely 'one-way-trip' during the period with the most informative data or the fact that northeast Pacific `r spp` are a single stock that exhibit movement throughout their range. In addition, the inclusion of new data from the same sources as the 2019 benchmark resulted in greater uncertainty around the derived quantities, a pattern which emerged as soon as natural mortality was freed and most pronounced in the slighly increased uncertainty around growth parameters.

Historical catches provide some information about the minimum stock size needed to have supported the observed time-series but there is less information about the upper bound on stock size. Likelihood profiles, parameter estimates, and general model behavior illustrate that small changes in many parameters can result in differing point estimates for management reference points, however the uncertainty about these estimates remains large unless leading model parameters, such as $M$ and $h$, are fixed. This uncertainty will remain until a more informative time-series and better quality demographic and biological information are accumulated for the stock, and potentially until a range wide northeast Pacific `r spp` analysis is available.

Uncertainty in the properties of current aging methods (both potential bias and imprecision), as well as relatively sparse fishery sampling, result in potentially noisy age data. Similarly, because `r spp` grow very rapidly and reach near-asymptotic length in their first decade of life, length-composition data were not particularly informative about historical patterns in recruitment. The patterns observed in historical `r spp` recruitment suggest that the stock trajectory (via shifts in recruitment strength) was linked to productivity regimes in the California Current. Uncertainty in future environmental conditions, changes in the timing, dynamics, and productivity of the California Current ecosystem, via climate change or cycles similar to the historical period, should be considered as a significant source of uncertainty in projections of stock status.

The WCGBT Survey was an excellent relative index of abundance over a broad demographic component of the `r spp` stock (although not the entire stock, as some of it occurs in deep water and was therefore unobserved).  This index, as well as stock assessments that better capture the dynamics of `r spp` across the NE Pacific, may inform future stock assessments about the scale of the `r spp` population relative to the catches being removed. The reduced survey effort (by 50\%) in 2019 and lack of survey at all in 2020 certainly reduced recent information available for this assessment. Having a complete survey in future years is of great importance for future iterations of this assessment.

A small set of sensitivity analyses were chosen to provide more information about potential information in the data, and potentially conflicting signals among data sources.\begin{comment}relatively obvious questions for any stock assessment such as sensitivity to key structural choices, potential information in the data, and potentially conflicting signals among data sources. The order in which they are presented was not intended to reflect their importance; each run included here provided important information for developing or evaluating the base model and alternate states of nature.\end{comment}The results are by no means meant to be a comprehensive comparison of all possible aspects of model uncertainty, nor do they reflect even the full range of models considered in developing the base model.

\begin{enumerate}
\item Use of the McAllister-Ianelli (Harmonic Mean) data weighting method as an alternative to the Francis method (described above in Section \ref{dw} on Data Weighting).
\item Use of the 2019 Francis data weights in lieu of the tuned Francis weights (described above in Section \ref{dw} on Data Weighting).
\item Inclusion of the At-Sea Hake Observer Program data (landings and length compositions). Adding information about `r spp` abundance gained from the Pacific hake (\emph{Merluccius productus}) fishery did not lead to significant changes relative to the base model; this model used the same weights as was done for the sensitivity in the 2019 benchmark analysis (Figure \ref{fig:ASHOP}). A sensitivity run with the Francis weight for the At-Sea Hake length compositions set to 1 did not converge (results not shown).
\item Inclusion of the commercially-landed length compositions for the fixed-gear and trawl fleets, using the same Francis data weights as in the base model. The time series of spawning biomass and depletion for this sensitivity were similar to the base model, though there were small changes in the magnitude of recruitment events in the 1970s (Figure \ref{fig:commL}). While we did not retune the Francis weights after the inclusion of this data, the tuning algorithm suggested to downweight the fixed-gear compositions by about half, and to upweight the trawl-gear lengths by a factor of about two. This is consistent with the trawl fishery, which samples more of the population, containing more information about incoming recruits.
\item Estimating a single parameter for natural mortality ($M$).  In the base model, $M$ was estimated at 0.0726 for females and 0.0605 for males; the single-parameter model estimated it to be lower at 0.052. The estimated unfished spawning biomass, while within the uncertainty bounds of the base model, was below the base model value. The sensitivity model reduced the size of large recruitment events and suggested the stock to be just barely above $B_{40\%}$ in 2020 (Figure \ref{fig:singleM}).
\item Implementing asymptotic age-based selectivity for the WCGBT Survey. This reduced overall model uncertainty (principally through reducing the standard deviation of $R_0$) and also reduced the size of large recruitment events.  This model had a higher overall log-likelihood than the base model, and did a poorer job of fitting the length compositions from that survey, particularly in the last year of data (Figure \ref{fig:asympNWFSC}).
\item Removing the index of sea surface height. A model run with the SSH data removed was identical to the base model in terms of depletion from the late 1970s onward, and shifted the large recruitment event backward by 3 years. (Figure \ref{fig:nossh}).
\item We also explored various parameter phasings, which had little impact on the base model.
\end{enumerate}


\begin{comment}
The following model changes to data or parameter estimation had little impact on
the base model.
\begin{enumerate}
\item Parameter phasing.
\item Estimating autocorrelation in the recruitment deviations.
\end{enumerate}

Removing the sea level index of recruitment from the base model resulted in a stock trajectory that was highly similar to, and within, the range of uncertanty estimated in the base model (Table \ref{tab:sensitivity_nosealevel}; Figure \ref{sensitivity_nosealevel}).   Small differences in model estimates were driven by differences in recruitment estimates, largely those prior to 1980 before age-composition data are available (Figure \ref{sensitivity_nosealevel})
Results from a model run using the 2015 selectivity patterns were within the range of uncertainty estimated in the base model, although estimates of both unfished spawing biomass and stock status were lower (Table \ref{tab:sensitivity_nosealevel}; Figure \ref{sensitivity_nosealevel_rec}).
Removing the WCGBT Survey index resulted in greater estimates of uncertainty around time series of spawning biomass and stock status, but more optimistic estimates of stock size and status at the end of the time series (Table \ref{tab:sensitivity_nosealevel}; Figure \ref{sensitivity_nosealevel}). 

A model runs implementing the following changes were largely withing the range of uncertainty estimated in the base model: 1) adding a hake bycatch fleet, 2) beginning the model in 1970 (a STAR panel request), and 3) estimating a single sex combined value for natural mortality (Table \ref{tab:ashop} and Figure \ref{fig:ashop}). 

In the post-STAR model adding a hake bycatch fleet resulted in a lower estimate of unfished spawning biomass. This difference is likely due to the removal of all other length composition data except for the WCGBTS data and, in this sensitivity run, the hake discard length compositions. Young (i.e., age-0) fish are caught in this mid-water trawl fishery as bycatch and it was hypothesized that including `r spp` lengths sampled by the hake fishery would be informative about recruitment. However, the time series does not appear to be long enough relative to the modeled period to be informative and the ongoing WCGBT Survey samples age-0 `r spp`. 

 Estimates of unfished spawning biomass and stock status in the single M run were lower than the base model value, while the estimate of unfished biomass and stock status were higher in the run that began during 1970.The scale of the estimated unfished spawning biomass is uncertain. To get a ball-park estimate of the scale of the northeast Pacific `r spp` population, conditioned on the `r coast` Coast assumptions and biology, a model run adding all northeast Pacific landings was completed. This model run suggests a northeast Pacific `r spp` population that follows a similar stock trajectory and results in a similar stock status compared to the base model, but that unfished spawning biomass could have ranged from about 250,000 mt to about 1,500,000 mt (Table \ref{tab:sensitivity_nepacific}; Figure \ref{sensitivity_nepacific}). This model sensitivity addresses, in a limited way, the long standing request for investigations into transboundary stock issues.

Models with a range of specifications for the age that defines the beginning of the plus group for the age data agreed regarding the strong increase in the spawning biomass during the 1960s to mid-1970s, followed by stock declines until recent years (Figures \ref{modelcomparison_dataage_ssb} and \ref{modelcomparison_dataage_rec}). The ages largely agreed regarding a period of high recruitments that drive this stock increase. This pattern was evident but less extreme in the 2011 and 2015 models due to the plus group being set at age 35. A similar pattern in spawning biomass trends was present in the AFSC stock assessment {@hanselman2018}. Extending the plus group to age 50 allows for fish aged 35 to 50 to better inform what historical recruitment may have been during periods that previous models assumed there was no information regarding recruitment as well as provides the potential to track truncation or expansion of the 'old growth' population age structure due to changes in fishing pressure or recruitment. 
In aggregate, these sensitivity analyses reflect the uncertainty in absolute stock size in this `r spp` assessment. Hopefully, they also provide a basis for future investigations, as well as a method for prioritizing potential research studies.
\end{comment}
## Retrospective analysis

A retrospective analysis was conducted by running the base model with
data removed for the past 5 years. All retrospective model runs fall within the uncertainty estimates from the base model. There was limited evidence of a retrospective pattern in estimates of spawning biomass and stock status, such that the view of the stock becomes more pessimistic as data are removed (Figures \ref{fig:retroSB} and \ref{fig:retroDEP}). The retrospective pattern in stock status is largely driven by the relative amount of data available to inform the estimates of some of the largest recruitments observed for `r spp` during 2013 and 2016 (Figure \ref{fig:retroREC}). 

## Historical analysis

Estimates of the current stock size and relative depletion were highly consistent with prior stock assessments, particularly from the 1970s forward, the period of time with good data for sablefish (Figure \ref{fig:histassess}). Estimates of stock size prior to the mid-1970s are greater in the 2005 and 2007 assessments, however there were limited data to inform the pre-1970 model period. Notably, depletion estimates for retrospective runs which truncate the series to 2019 or earlier are within the narrower uncertainty bounds from the 2019 benchmark, while the perception of the stock as increasing is consistent across all models which include the new data and timeblocking structure (Figures \ref{fig:retrobenchDEP} and \ref{fig:retrobenchREC}).

## Likelihood profiles

Likelihood profiles were used to elucidate conflicting information among various data sources, to determine how asymmetric the likelihood surfaces surrounding point estimates may be, and to provide an additional evaluation of how precisely parameters are being estimated. Likelihood profiles were completed for three key model parameters: female $M$, unexploited equilibrium recruitment ($R_0$), and $h$. For a single parameter (loosely interpreting an iteratively re-weighted stock assessment objective function in terms of true likelihood) an increase in negative log-likelihood of more than two units indicates a statistically significant degradation in fit.

Female $M$ (male natural mortality rate estimates are highly correlated with female mortality, so it is not included in this discussion) was found to be moderately informed across a relatively wide range of values. Data from the discard appears to be the most influential for this parameter. Differences in total negative log likelihood was less than two across approximately 0.060-0.095 for female `r spp` $M$ Figures \ref{fig:sens_m_like}). However, this is not a trivial parameter range and the assessment results vary considerably among these values in absolute scale (Figure \ref{fig:sens_m_ssb}).

Unexploited equilibrium recruitment ($R_0$) was found to have similar likelihoods over 9.2-10.4, values which led to a broad range of stock sizes (Figures \ref{fig:sens_R0_like}-\ref{fig:sens_R0_ssb}). The range of values explored led to little differences in the current level of depletion the stock is facing but large differences in depletion from 1935 to 1970 where there is little information during a period with fishing (Figure \ref{fig:sens_R0_depl}).

In the base model, $h$ is fixed at `r model$parameters[grep("steep", model$parameters$Label), "Value"]`, making it an important profile to evaluate as its uncertainty is not explicitly included in the base-model results. In 2011, the maximum likelihood estimate for $h$ was 0.2, which implies zero surplus production, which is biologically implausible. This assessment found essentially equal support in the data over a broad range of explored values (Figure \ref{fig:profile_like_h}). Most of the values included in the profile led to similar trajectories of spawning biomass (Figure \ref{fig:profile_ssb_h}).\begin{comment} The relative strengths of recent cohorts were also not strongly influenced by the value for $h$ (Figure \ref{sens_h_rec}), and the relative depletion level is quite robust as well (Figure \ref{sens_h_depl}). Uncertainty from $h$ was well inside the global estimation uncertainty captured via the asymptotic intervals about the maximum likelihood estimates.
\end{comment}

In aggregate, these profiles explain why the asymptotic uncertainty about historical and current stock size is so broad and underscore the lack of information in the data regarding scale for this stock assessment.

# Reference points{#rp}
\textrefpt

\textexplstatus
The phase plot shows the interaction of fishing intensity and biomass targets (Figure \ref{fig:SPR4_phase}).

# Harvest projections and decision tables


\textprojections 
 
 
# Regional management considerations

Recent `r spp` management has relied upon apportionment of the ACL north and south of $36^\circ$ N latitude using the average estimated differences in biomass from the WCGBT Survey. This historical management line corresponds with a recent data-driven analysis of `r spp` growth that suggests a difference in growth rates north and south of $36^\circ$ N latitude (@kapur2020). The estimates represent the relative distribution of the `r spp` population observed by the survey, not the entire population. Additionally, it is likely that fish from more northerly regions are migrating into U.S. West Coast waters (pers. comm., L. Rogers), which may bias the survey estimates of the distribution of fish in each region. Thus, these results should be interpreted with caution.

```{r, include = FALSE}
split36 <- read.table("table_for_ss3_correctedNS.csv", comment = "^", sep = ",", header = TRUE) %>%
  filter(Year<2020)
splits <- aggregate(Estimate_metric_tons ~ Year, 
  data = split36[split36$Fleet %in% c("North", "South"), ], prop.table)
colnames(splits)[2] <- ""
colnames(splits[, 2]) <- c("South", "North")
splits.use <- apply(splits[, 2], 2, mean)
```
The average survey biomass, from 2003 to `r last(model$cpue$Yr)`, that has been distributed south of $36^\circ$N, is `r round(splits.use["South"]*100, 2)`%. The average survey biomass, from 2003 to 2019, that has been distributed north of $36^\circ$N, is `r round(splits.use["North"]*100, 2)`%. The 2015 and 2019 assessments estimated that 26.2% and 26.3% of the biomass was found south of Point Conception and 73.8% and 73.7% of the biomass was found to the north, respectively. The estimates from the WCGBT Survey show that the spatial distribution of `r spp` along the U.S. West Coast appears to be relatively stable, particularly from 2008 to 2014 (Figure \ref{fig:WCGBTindexest}).



